---
title: "Final Report"
author: "Nan Li"
date: "June 16, 2019"
output:
  pdf_document: 
  html_document:
    css: faded.css
---
## Data

The `gapminder` data set includes 6 major variables that describes life expentency depending upon factors such as region, population, GDP, and some others. Below are the major variables: 


* **life** - life expectancy
* **income** - gdp per capita
* **year** - the range from 1800 to 2015
* **county** - countries in the world
* **region** - 6 regions includes all contries
* **population** - census data collected about every 10 years

## TWO major questions that I listed:
1. What is the difference in GDP numbers across various regions? And what are the rankings? 
2. How is the relationship between GDP and life expectancy?

 
## Data summary
```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
rm(list = ls())
myData <- read.csv("file:///C:/Users/nanl/Desktop/HW/ANLY506/gapminder.csv", header=TRUE)
head(myData)
tail(myData) 
names(myData)
summary(myData)
nObs = nrow(myData)
nVars = ncol(myData)
year2015 <- filter(myData, Year == 2015)
```

1. There are `r nObs` observations. 
2. Thee are `r nVars` variables. 
3. Type of variables: "Country" is `r class(myData$Country)`, "Year" is `r class(myData$Year)`, "life" is `r class(myData$life)`, "population" is `r class(myData$population)`, "income" is `r class(myData$income)` and "region" is `r class(myData$region)`.
4. Data range: Range of "Year" is `r range(myData$Year)`. Range of "life" is `r range(myData$life)`. Range of "income" is `r range(myData$income, na.rm = TRUE)`.
5. Data wrangling: The avaerage life expectancy in year 2015 is `r mean(year2015$life, na.rm = TRUE)`.
6. Missing values: in the data file, there are some missing values about population data. So I used 'fill' method, and removed the rows with empty income values. I also converted the type of population from factor to numeric, so that is easier to deal with. 

```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
gm = filter(myData, !is.na(myData$income))
class(gm$population)
gm$population = as.numeric(as.character(gm$population))
class(gm$population)
gm <- gm %>% fill(population)
```

## Exploratory Plot for Question_1
```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
ggplot(gm, aes(region, income, color = region)) + geom_boxplot() + labs(title="GDP Per Region", y = "GDP", x = "Region")
```

Figure. 1 shows the GDP numbers are different across regions. 

## Exploratory Plot for Question_2
```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
ggplot(data = gm, mapping = aes(x = income, y = life, color = region)) +
  geom_point(alpha = 0.5) +
  scale_x_log10()+
  scale_color_manual(values = c("red","yellow", "orange","green","blue","purple")) + 
  labs(title="The relation between GDP and Life Expenctancy", y = "Life Expectancy", x = "GDP")
```

Figure. 2 shows the relationship between GDP and Life Expenctancy, which is positive. 


## Answers to the TWO questions:
1. From Figure. 1, we can clearly see that GDP numbers are different across different regions. And the ranking are as following: No.1 Europe & Central Asia; No.2 East Asia & Pacific; No.3 Europe & Central Asia; No.4 America; No.5 Sub-Saharan Africa; No.6 South Asia.
2. From Figure. 2, we can tell a positive relationship between GDP and life expectancy. For example, region Middle East & North Africa has high GDP/outincome, and they also have high life expectancy (average is in the >60 range)

